---
#title: "Analyzing Stock Market Data with R - Trend Analysis Ehanced tweaks""
#output: html_document
#description: working through the design of a trading system enhanced.
---

Trading With The Trend, page 2 expanding

```{r , echo=TRUE, fig.width=10, fig.height=7}
library('quantmod')
getSymbols(c('SPY', 'AMZN', '^GSPC'), src='yahoo')
# install.packages('binhf')
library('binhf')
library('TTR')
```

```{r , echo=TRUE, fig.width=10, fig.height=7}
SPY.EMA.10 <- EMA(SPY$SPY.Close, n=10 ) 
SPY.EMA.50 <- EMA(SPY$SPY.Close, n=50 ) 
SPY.EMA.200 <- EMA(SPY$SPY.Close, n=200 ) 
SPY.Fast.Diff <- SPY.EMA.10 - SPY.EMA.50
SPY.Slow.Diff <- SPY.EMA.50 - SPY.EMA.200
```

```{r , echo=TRUE, fig.width=10, fig.height=7}
SPY.Long_Trades <- ifelse(
      SPY.Slow.Diff > 0 &
      SPY.Fast.Diff > 0 & 
      shift(v=as.numeric(SPY.Fast.Diff), places=1,dir = 'right') < 0,
      SPY$SPY.Close, NA)

SPY.Short_Trades <- ifelse(
      SPY.Slow.Diff < 0 &
      SPY.Fast.Diff < 0 & 
      shift(v=as.numeric(SPY.Fast.Diff), places=1,dir = 'right') > 0,
      SPY$SPY.Close, NA)

plot(SPY$SPY.Close)

points(SPY.Long_Trades, col='blue', cex=1.5, pch=18)
points(SPY.Short_Trades, col='red', cex=1.5, pch=18)

# kick out pdf of last plot
dev.copy(pdf, "/Users/markloessi/InvestingCode/R_Investing/SPY-SPY-CLOSE-trendInOut1050200.pdf", width=10, height=7)
dev.off()
```

Slowing down the moving averages might help us see things better, the above has to many poor trades being identified. 

Let’s see how trying different moving-average periods affect the entry system. Instead of the 10,50,200 series, let’s try both indexes on 20, 100, 300 on the S&P 500:

```{r , echo=TRUE, fig.width=10, fig.height=7}
SPY.EMA.FastS <- EMA(SPY$SPY.Close, n=20 ) 
SPY.EMA.MediumS <- EMA(SPY$SPY.Close, n=100 ) 
SPY.EMA.SlowS <- EMA(SPY$SPY.Close, n=300 ) 
SPY.Fast.DiffS <- SPY.EMA.FastS - SPY.EMA.MediumS
SPY.Slow.DiffS <- SPY.EMA.MediumS - SPY.EMA.SlowS
```

```{r , echo=TRUE, fig.width=10, fig.height=7}
# look for long entries
SPY.Long_TradesS <- ifelse(
        SPY.Slow.DiffS  > 0 &
                SPY.Fast.DiffS  > 0 &
                shift(v=as.numeric(SPY.Fast.DiffS), places=1, dir="right") < 0,
                SPY$SPY.Close, NA)

# look for long exits (same thing but inverse signts)
SPY.Short_TradesS <- ifelse(
        SPY.Slow.DiffS  < 0 &
                SPY.Fast.DiffS  < 0 &
                shift(v=as.numeric(SPY.Fast.DiffS), places=1, dir="right") > 0,
                SPY$SPY.Close, NA)

plot(SPY$SPY.Close)

## Warning in plot.xts(SPY): only the univariate series will be plotted
points(SPY.Long_TradesS, col='blue', cex=1.5, pch=18)
points(SPY.Short_TradesS, col='red', cex=1.5, pch=18)

# kick out pdf of last plot
dev.copy(pdf, "/Users/markloessi/InvestingCode/R_Investing/SPY-SPY-CLOSE-trendInOut10100300.pdf", width=10, height=7)
dev.off()
```

Adding extra rules

Let’ try adding an interesting rule twist. Instead of entering when the slower indicator (red colored difference of the slower moving averages) is either above zero for longs and below zero for shorts, let’s enter when its moving in the trade’s direction.

New rules are:

    if no position: red-1 < red and blue-1 < 0 and blue > 0 go long
    if long: blue < 0 exit long
    
    if no position: red-1 > red and blue-1 > 0 and blue < 0 go short
    if short: blue > 0 exit short

```{r , echo=TRUE, fig.width=10, fig.height=7}
# duplicate from above only so we can see it
SPY.EMA.FastS <- EMA(SPY$SPY.Close, n=20 ) 
SPY.EMA.MediumS <- EMA(SPY$SPY.Close, n=100 ) 
SPY.EMA.SlowS <- EMA(SPY$SPY.Close, n=300 ) 
SPY.Fast.DiffS <- SPY.EMA.FastS - SPY.EMA.MediumS
SPY.Slow.DiffS <- SPY.EMA.MediumS - SPY.EMA.SlowS
```
A for adjusted theory, aka we've enhanced the logn and short thinking.
```{r , echo=TRUE, fig.width=10, fig.height=7}
# look for long entries
SPY.Long_TradesSA <- ifelse(
        SPY.Slow.DiffS  > # when the slow diff is greater than the previous slow dif
        shift(v=as.numeric(SPY.Slow.DiffS), places=1, dir="right") &
        SPY.Fast.DiffS  > 0 & # and the fast diff is greater than Zero
        shift(v=as.numeric(SPY.Fast.DiffS), places=1, dir="right") < 0, # and previous fast is less than Zero
                SPY$SPY.Close, NA) 

# look for long exits (same thing but inverse signts) ala the shorts
SPY.Short_TradesSA <- ifelse(
        SPY.Slow.DiffS  < 
        shift(v=as.numeric(SPY.Slow.DiffS), places=1, dir="right") &       
        SPY.Fast.DiffS  < 0 &
        shift(v=as.numeric(SPY.Fast.DiffS), places=1, dir="right") > 0,
        SPY$SPY.Close, NA)

plot(SPY$SPY.Close)

## Warning in plot.xts(SPY): only the univariate series will be plotted
points(SPY.Long_TradesSA, col='blue', cex=1.5, pch=18)
points(SPY.Short_TradesSA, col='red', cex=1.5, pch=18)

# kick out pdf of last plot
dev.copy(pdf, "/Users/markloessi/InvestingCode/R_Investing/SPY-SPY-CLOSE-trendInOut20100300Enh.pdf", width=10, height=7)
dev.off()
```

```{r , echo=TRUE, fig.width=10, fig.height=7}
chartSeries(SPY, theme="white",
            TA="addEMA(50, col='black');addEMA(300, col='blue')")
addTA(SPY.Fast.DiffS, col='blue', type='h',legend="20-100 MA")
addTA(SPY.Fast.DiffS, col='red', type='h',legend="100-300 MA")
```

What we've learned here is that all the choppy goings on here is overwhelming the trending system as what we are seeing on our latest sysem (reproduced below) with the shorter EMA's is really not producing something that we'd want to use on this stock. 

Lets go back to 10 - 50 - 200 for this display.
```{r , echo=TRUE, fig.width=10, fig.height=7}
SPY.EMA.Fast <- EMA(SPY$SPY.Close, n=10 ) 
SPY.EMA.Medium <- EMA(SPY$SPY.Close, n=50 ) 
SPY.EMA.Slow <- EMA(SPY$SPY.Close, n=200 ) 
SPY.Fast.Diff <- SPY.EMA.Fast - SPY.EMA.Medium
SPY.Slow.Diff <- SPY.EMA.Medium - SPY.EMA.Slow
```

```{r , echo=TRUE, fig.width=10, fig.height=7}
# look for long entries
SPY.Long_TradesA <- ifelse(
        SPY.Slow.Diff  > # when the slow diff is greater than the previous slow dif
        shift(v=as.numeric(SPY.Slow.Diff), places=1, dir="right") &
        SPY.Fast.Diff  > 0 & # and the fast diff is greater than Zero
        shift(v=as.numeric(SPY.Fast.Diff), places=1, dir="right") < 0, # and previous fast is less than Zero
                SPY$SPY.Close, NA) 

# look for long exits (same thing but inverse signts) ala the shorts
SPY.Short_TradesA <- ifelse(
        SPY.Slow.Diff  < 
        shift(v=as.numeric(SPY.Slow.Diff), places=1, dir="right") &       
        SPY.Fast.Diff  < 0 &
        shift(v=as.numeric(SPY.Fast.Diff), places=1, dir="right") > 0,
        SPY$SPY.Close, NA)

plot(SPY$SPY.Close)

## Warning in plot.xts(SPY): only the univariate series will be plotted
points(SPY.Long_TradesA, col='blue', cex=1.5, pch=18)
points(SPY.Short_TradesA, col='red', cex=1.5, pch=18)

# kick out pdf of last plot
dev.copy(pdf, "/Users/markloessi/InvestingCode/R_Investing/SPY-SPY-CLOSE-trendInOut1050200Enh.pdf", width=10, height=7)
dev.off()
```





