---
title: "PracticalDataScience_AnalyzingStocks_InsightsCommonIndicators"
output: html_document
---

Insight from Common Indicators

```{r , echo=TRUE, fig.width=10, fig.height=7}
library('quantmod')
getSymbols(c('SPY', 'AMZN', 'QQQ'), src='yahoo')

# remove any NAs 
SPY <- SPY[!(rowSums(is.na(SPY))),]
AMZN <- AMZN[!(rowSums(is.na(AMZN))),]
QQQ <- QQQ[!(rowSums(is.na(QQQ))),]

# install.packages('binhf')
library('binhf')
library('TTR')
```

### ADX

The ADX is Welles Wilder’s Directional Movement Indicator. It is used by lots of people to determine if the market is trending or range bound.

Blue line interpretation; Blue above 20 is a trend
Red is the bearish indicator
Green is bullish indicator
You read the Red to Green as what's going on, so to speak and the Blue for strength of trend

Example;
Red over Green and Blue above 20 means strong bearish trend. 

```{r , echo=TRUE, fig.width=10, fig.height=7}
chartSeries(QQQ, theme="white", TA="addSMA(50, col='black');addSMA(200, col='blue');addADX(n = 14, maType='EMA', wilder=TRUE)", subset='2017::')
```

```{r , echo=TRUE, fig.width=10, fig.height=7}
chartSeries(SPY, theme="white", TA="addSMA(50, col='black');addSMA(200, col='blue');addADX(n = 14, maType='EMA', wilder=TRUE)", subset='2017::')
```

In a nutshell, Welles recommends using the ADX with a 14-day period. When the main blue line is above 20, it is considered a strong, trending market, when it is below, it is considered a weak one.

### Volume VWAP - directional indicator

As this evaluation is mostly using the closing price but it is important to note that there are a lot of other market variables available. You can design systems with the open price, the high or low, the difference between the open and close, etc. And there is also the volume.

This an important indicator. A falling stock price on rising volume or a rising stock on falling volume may mean the move is about to reverse. Whatever the reason for abnormal volume, it should be a warning to keep a vigilant eye on the stock.

There are plenty of indicators that include the volume price such as the Volume-weighted average price (VWAP). The VWAP is a guide more than a trading indicator as to where the market is trading compared to the volume adjusted price. It divides dollars traded by volume (see above link for more details). Pay attention to the zero line.

```{r , echo=TRUE, fig.width=10, fig.height=7}
SPY.VWAP.Slow <- VWAP(price=SPY$SPY.Close, volume=SPY$SPY.Volume, n=100)
SPY.VWAP.Fast <- VWAP(price=SPY$SPY.Close, volume=SPY$SPY.Volume, n=20)
SPY.VWAP.Diff <- SPY.VWAP.Fast - SPY.VWAP.Slow
```

```{r , echo=TRUE, fig.width=10, fig.height=7}
chartSeries(SPY, theme="white", subset='2017::',
            TA="addVo();addTA(SPY.VWAP.Slow, on=1, col='red');
              addTA(SPY.VWAP.Fast, on=1, col='blue');
              addTA(SPY.VWAP.Diff, col='blue');
              addADX(n = 14, maType='EMA', wilder=TRUE)")

# kick out pdf of last plot
dev.copy(pdf, "/Users/markloessi/InvestingCode/R_Investing/SPY-SPY-CLOSE-VWAP_ADX.pdf", width=10, height=7)
dev.off()
```

```{r , echo=TRUE, fig.width=10, fig.height=7}
QQQ.VWAP.Slow <- VWAP(price=QQQ$QQQ.Close, volume=QQQ$QQQ.Volume, n=100)
QQQ.VWAP.Fast <- VWAP(price=QQQ$QQQ.Close, volume=QQQ$QQQ.Volume, n=20)
QQQ.VWAP.Diff <- QQQ.VWAP.Fast - QQQ.VWAP.Slow
```

```{r , echo=TRUE, fig.width=10, fig.height=7}
chartSeries(QQQ, theme="white", subset='2017::',
            TA="addVo();addTA(QQQ.VWAP.Slow, on=1, col='red');
              addTA(QQQ.VWAP.Fast, on=1, col='blue');
              addTA(QQQ.VWAP.Diff, col='blue');
              addADX(n = 14, maType='EMA', wilder=TRUE)")

# kick out pdf of last plot
dev.copy(pdf, "/Users/markloessi/InvestingCode/R_Investing/QQQ-QQQ-CLOSE-VWAP_ADX.pdf", width=10, height=7)
dev.off()
```

```{r , echo=TRUE, fig.width=10, fig.height=7}
AMZN.VWAP.Slow <- VWAP(price=AMZN$AMZN.Close, volume=AMZN$AMZN.Volume, n=100)
AMZN.VWAP.Fast <- VWAP(price=AMZN$AMZN.Close, volume=AMZN$AMZN.Volume, n=20)
AMZN.VWAP.Diff <- AMZN.VWAP.Fast - AMZN.VWAP.Slow
```

```{r , echo=TRUE, fig.width=10, fig.height=7}
chartSeries(AMZN, theme="white", subset='2017::',
            TA="addVo();addTA(AMZN.VWAP.Slow, on=1, col='red');
              addTA(AMZN.VWAP.Fast, on=1, col='blue');
              addTA(AMZN.VWAP.Diff, col='blue');
              addADX(n = 14, maType='EMA', wilder=TRUE)")

# kick out pdf of last plot
dev.copy(pdf, "/Users/markloessi/InvestingCode/R_Investing/AMZN-AMZN-CLOSE-VWAP_ADX.pdf", width=10, height=7)
dev.off()
```

### Make a trading system

```{r , echo=TRUE, fig.width=10, fig.height=7}
AMZN.VWAP.Slow <- VWAP(price=AMZN$AMZN.Close, volume=AMZN$AMZN.Volume, n=100)
AMZN.VWAP.Fast <- VWAP(price=AMZN$AMZN.Close, volume=AMZN$AMZN.Volume, n=20)
AMZN.VWAP.Diff <- AMZN.VWAP.Fast - AMZN.VWAP.Slow

AMZN.ADX.14 <- ADX(AMZN,n=14)
```

Adjust your VWAP and ADX numbers to correspond to something meaningful for your stock. 
```{r , echo=TRUE, fig.width=10, fig.height=7}
# look for long entries
AMZN.Long_Trades <- ifelse(
        AMZN.ADX.14$ADX > 20 &
        AMZN.VWAP.Diff > 100, AMZN$AMZN.Close, NA)

# look for long entries
AMZN.Short_Trades <- ifelse(
        AMZN.ADX.14$ADX > 20 &
        AMZN.VWAP.Diff < -100, AMZN$AMZN.Close, NA)

plot(AMZN$AMZN.Close, subset='2017::')

## Warning in plot.xts(QQQ): only the univariate series will be plotted
points(AMZN.Long_Trades, col='blue', cex=1, pch=18)
points(AMZN.Short_Trades, col='red', cex=1, pch=18)
```

### AMZN Kitchen Sink

```{r , echo=TRUE, fig.width=10, fig.height=20}
AMZN.EMA.22<- EMA(AMZN$AMZN.Close, n=22)
AMZN.EMA.50<- EMA(AMZN$AMZN.Close, n=50)
AMZN.EMA.100<- EMA(AMZN$AMZN.Close, n=100)
AMZN.EMA.200<- EMA(AMZN$AMZN.Close, n=200)
AMZN.Fast.Diff <- AMZN.EMA.22 - AMZN.EMA.50
AMZN.Slow.Diff <- AMZN.EMA.50 - AMZN.EMA.200
chartSeries(AMZN, theme="white", subset='2018::',
            TA="addVo();
              addWPR(n=14);
              addMACD(type='EMA');
              addRSI(n = 14, maType='EMA', wilder=TRUE);
              addSMI(n = 13, slow = 25, fast = 2, signal = 9, ma.type='EMA'); #hmm... Stochastic Oscillator
              addADX(n = 14, maType='EMA', wilder=TRUE);
              addCCI();
              addBBands();
              addROC(n=200,col = 'blue');addROC(n=50,col = 'red');addROC(n=22,col = 'green') # rate of change
            ")
addTA(AMZN.EMA.22, on=1, col = "green")
addTA(AMZN.EMA.50, on=1, col = "yellow")
addTA(AMZN.EMA.100, on=1, col = "blue")
addTA(AMZN.EMA.200, on=1, col = "red")
addTA(AMZN.Slow.Diff, col='red', type='h',legend="Slow Diff 50-200 MA used for in-out of market")
addTA(AMZN.Fast.Diff, col='blue', type='h',legend="Fast Diff 22-50 MA used for in-out of market")

# kick out pdf of last plot
dev.copy(pdf, "/Users/markloessi/InvestingCode/R_Investing/AMZN-KitchenSink.pdf", width=10, height=20)
dev.off()
```
